---
title: "Understanding the COVID-19 pandemic through data"
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    css: R-shiny-style-changes.css
---
  
```{r setup, include=FALSE}
source("styles.R")
source("00_setup.R")

library(pins)
library(RColorBrewer)
library(data.table)


my_key <- Sys.getenv("connect_key")
my_server <- "http://localhost:3939/"

# UNCOMMENT FOR LOCAL USE with EXTERNAL server:
# my_key <- Sys.getenv("connect_key_ext")
# my_server <- "http://w0lxopshyprd1b.worldbank.org:3939/" # External server

# UNCOMMENT FOR LOCAL USE with INTERNAL server:
# my_key <- Sys.getenv("connect_key_int")
# my_server <- "http://w0lxpjekins05.worldbank.org:3939/" # Internal server

board_register_rsconnect(key    = my_key,
                         server = my_server)

#--------- Data from server
dfm            <- pin_get("dfm", board = "rsconnect")
ccc            <- pin_get("ccc", board = "rsconnect")
fullcd2        <- pin_get("fullcd2", board = "rsconnect")
treemap        <- pin_get("treemap_data", board = "rsconnect")
wld_data       <- pin_get("wld_data", board = "rsconnect")
indicator_list <- pin_get("indicator_list", board = "rsconnect")
new_c          <- pin_get("new_c", board = "rsconnect")
new_wl         <- pin_get("new_wl", board = "rsconnect")
#cg_data       <- pin_get("cg_data", board = "rsconnect")
age_pop        <- pin_get("age_pop", board = "rsconnect")
age_pop        <- age_pop$age_pop
covid          <- pin_get("covid", board = "rsconnect")
health         <- pin_get("health", board = "rsconnect")
water          <- pin_get("water", board = "rsconnect")
df             <- pin_get("dfsev", board = "rsconnect")


#--------- covid data
covid          <- covid$covid
covid          <- c( "COVID-19 Cases: Case Fatality Rate",
                     "COVID-19 Cases: Confirmed",
                     "COVID-19 Cases: Confirmed (log scale)",
                     "COVID-19 Cases: Confirmed (per 1,000 people)",
                     "COVID-19 Cases: Confirmed (per 1,000 people, log scale)",
                     "COVID-19 Cases: Deaths",
                     "COVID-19 Cases: Deaths (log scale)",
                     "COVID-19 Cases: Deaths (per 1,000 people)",
                     "COVID-19 Cases: Deaths (per 1,000 people, log scale)",
                     "COVID-19 Cases: New Confirmed cases (Daily)",
                     "COVID-19 Cases: New Deaths (Daily)")
covidlog       <- covid[stringr::str_detect(covid, "log scale")]
covidbubble    <- covid[!stringr::str_detect(covid, "log scale")]
covid_treemap  <- covid[!stringr::str_detect(covid, paste(c("log scale", "Fatality", "per 1,000 people"),collapse = '|'))]


#--------- Aux data
health         <- health$health
water          <- water$water

my_date  <- max(wld_data$date)
my_date2 <- format(as.Date(my_date), format = "%b %d")
my_date2 <- as.character(my_date2)
today    <- as.Date(my_date) + 1

#--------- filtered cases
covid <- grep("(Confirmed|Deaths)", covid, value = TRUE)


#------country palette
#This has to be done here and not in utils because Ineed the `fullcd2` dataframe

uniq_cty <- unique(fullcd2$`Short Name`)

cty_color <-  setNames(
  rep(palette, ceiling(length(uniq_cty)/length(palette)))[1:length(uniq_cty)],
  uniq_cty
)


income <- c("Income level")

#--------- Find countries that are doing really bad

grcd <- fullcd2 %>% 
  select(`Short Name`, date, variable, value) %>% 
  filter(!is.na(`Short Name`)) %>% 
  as.data.table()


wst <- grcd[
  order(`Short Name`, variable, date)
     ][,
       # growth of each variable
       `:=`(gr = value/shift(value, 1, type = "lag") - 1,
            df = value - shift(value, 1, type = "lag") ),
       by = .(`Short Name`, variable)
     ][,
       # Rolling mean
       `:=`(rmg = frollmean(gr, 7, align =  "right"),  # rolling mean of growth
            rmd = frollmean(df, 7, align =  "right")), # rolling mean of diff
       by = .(`Short Name`, variable)
     ][
       # Select only one variable
       variable == "COVID-19 Cases: Deaths" & !is.na(rmg),
       # variable == "COVID-19 Cases: Confirmed",
       
       # select mrv of each group
       .SD[.N],
       by = .(`Short Name`)
     ][
       order(-value)
     ][
       1:20
     ][
       order(-rmd)  # Let's use the rolling mean of diff to include the US
     ][
       1:6,
       `Short Name`
     ]
rm(grcd)


```


Indicators
=====================================
  
Sidebar { .sidebar data-width=250}
-----------------------------------------------------------------------
Last Updated on: `r today`

```{r}

selectInput("variable",
            "Indicator Y-axis:",
            choices = list(
              `Age & Population` = age_pop,
              `COVID-19` = covid,
              Health = health,
              `Water & Sanitation` = water),  
            selected = "COVID-19 Cases: Deaths"
)    

selectInput("variable2",
            "Indicator X-axis:",
            choices = list(
              `Age & Population` = age_pop,
              `COVID-19` = covid,
              Health = health,
              `Water & Sanitation` = water),                
            selected = "Population ages 65 and above (% of total)"
) 

#selectInput("size",
#            "Bubble size:",
#            choices = covidbubble,                
#            selected = "COVID-19 Cases: Confirmed"
#) 

radioButtons("color",
             "Bubble color (click on legend items to exclude selected category):",
             choices = c("Income Group", "Region"),  
             selected = "Income Group"
) 

selectizeInput("Country Name",
               "Country (compare countries over time):",
               c(unique(as.character(dfm$`Short Name`))),
               multiple = TRUE, 
               options = list(maxItems = 12),
               selected = wst
)
```

Row  {data-height=130}
-----------------------------------------------------------------------
### World
```{r}
renderValueBox({
  con <- wld_data %>%
    filter(wld_data$date == max(wld_data$date)) %>%
    pull(`COVID-19 cases: Confirmed`) 
  con <- format(con, big.mark=",")
  de <- HTML(paste0("Total cases:<br/>", con))
  valueBox(tags$p(de, style = "font-size: 80%;")
           #, 
           #icon = "fa-users"
  )
})
```

### World
```{r}
renderValueBox({
  con <- new_wl %>% 
    filter(status == "Confirmed") %>% 
    pull(con) %>% 
    format(big.mark = ",")
  
  de <- HTML(paste0("New cases (", my_date2, "):<br/>", con))
  #de <- str_wrap(de, 15)
  valueBox(tags$p(de, style = "font-size: 80%;")
           #, 
           #icon = "fa-users"
  )
})
```

### World
```{r}
renderValueBox({
  con <- wld_data %>%
    filter(wld_data$date == max(wld_data$date)) %>%
    pull(`COVID-19 cases: Deaths`) %>%
    format(big.mark = ",")
  
  de <- HTML(paste0("Total deaths:<br/>", con))
  valueBox(tags$p(de, style = "font-size: 80%;")
           #, 
           #icon = "fa-users"
  )
})
```

Row  {data-height=500}
-----------------------------------------------------------------------
  
### Explore relationships
```{r}
renderPlotly ({
  df2 <- dfm[dfm$date == 2019 & dfm$variable %in% input$variable,]
  df3 <- dfm[dfm$date == 2019 & dfm$variable %in% input$variable2, c(1, 6)]
  dfs <- dfm[dfm$date == 2019 &
          dfm$variable %in% "COVID-19 Cases: Deaths", c(1, 6)]
  
  dfs <- dfs %>%
    rename("siz" = "mrv")
  
  df3 <- df3 %>%
    rename("mrv2" = "mrv")
  
  df4 <- merge(df2, df3, by = c("iso3c"))
  df4 <- merge(df4, dfs, by = c("iso3c"))
  
  m <- df4[df4$`Short Name` %in% input$`Country Name`, ]
  a <- list(
    x = m$mrv2,
    y = m$mrv,
    text = m$`Short Name`,
    xref = "x",
    yref = "y",
    showarrow = TRUE,
    arrowhead = 0,
    ax = 20,
    ay = -40,
    arrowcolor = toRGB("grey80"),
    font = list(color = "dark grey", size = 14)
  )
  
  
  texty <- paste("</br>",
                 df4$`Short Name`,
                 "</br>X: ",
                 round(df4$mrv2, 1),
                 "</br>Y: ",
                 round(df4$mrv, 1))
  
  
  df4$`Income Group` <- ordered(
    df4$`Income Group`,
    levels = c(
      "Low income",
      "Lower middle income",
      "Upper middle income",
      "High income"
    )
  )
  
  if (input$color == "Income Group") {
    df4$colort <- df4$`Income Group`
  } else {
    df4$colort <- df4$Region
  }
  
  if (input$color == "Income Group") {
    colorpal <- inc_color
  } else {
    colorpal <- brewer.pal(7, "Paired")
  }
  
  plot_ly(data = df4) %>%
    add_trace(
      x = ~ mrv2,
      y = ~ mrv,
      type = 'scatter',
      mode = 'markers',
      color =  ~ colort,
      colors = ~ colorpal,
      opacity = 0.9,
      size   = 15,
      #  sizes  = c(20, 100),
      hoverinfo = 'text',
      text = ~ texty
    ) %>%
    #add_annotations(text = df4$lab, textfont = list(size = 14),
    #         textposition = "top right") %>%
    layout(
      legend = list(
        x = 0,
        y = -0.3,
        orientation = 'h'
      ),
      annotations = a,
      xaxis = list(
        type = if_else(input$variable2 %in% covidlog, "log", "linear"),
        zeroline = FALSE,
        showline = FALSE,
        showgrid = TRUE,
        title = str_wrap((input$variable2), 40),
        range = c(min(df4$variable), max(df4$variable))
      ),
      yaxis = list(
        type = if_else(input$variable %in% covidlog, "log", "linear"),
        zeroline = FALSE,
        showline = FALSE,
        showgrid = TRUE,
        title = str_wrap(input$variable, 40),
        range = c(min(df4$variable), max(df4$variable))
      )
    )
    }
) # end of renderPlotly
```


### Compare countries over time
```{r}

uiOutput("dynamic")

output$dynamic <- renderUI({ 
  if (input$variable %in% covid){
    plotlyOutput("plot1")
  } else {
    plotlyOutput("plot2")
  }
})

output$plot1 <- renderPlotly({
  
  dft <- fullcd2[fullcd2$`Short Name` %in% input$`Country Name`
                 & fullcd2$variable %in% input$variable,]
  
  dft <- dft         %>%
    arrange(date)    %>% 
    group_by(Region) %>% 
    mutate(
      op = match(name, unique(name)), 
      opacity = (1-(op/max(op))+1/max(op)),
    ) %>% 
    ungroup() %>% 
    mutate(
      colors = rep(palette, ceil(nrow(.)/length(palette)))[1:nrow(.)]
    )  
  
  
  texty <- paste("</br>", dft$`Short Name`,
                 "</br>Y: ", round(dft$value, 2), 
                 "</br>Date: ", dft$date)
  
  plot_ly(data = dft,
          x    = ~date,
          y    = ~ value ) %>%
    add_trace(

      name      = ~ `Short Name`, 
      type      = 'scatter',
      mode      = 'lines+markers', 
      color     = ~`Short Name`,
      colors    = brewer.pal(12, "Paired"),
      hoverinfo = 'text',
      text      = ~ texty
    ) %>%
    layout(
      legend     = list(orientation = 'h'),
      xaxis      = list(
        title    = '',
        zeroline = FALSE,
        showline = FALSE,
        showgrid = TRUE
      ),
      yaxis      = list(type = if_else(input$variable %in% covidlog, "log", "linear"),
        title    = input$variable,
        zeroline = FALSE,
        showline = FALSE,
        showgrid = TRUE
      )
    )
  
})

output$plot2 <- renderPlotly({
  
  df1 <- dfm[dfm$`Short Name` %in% input$`Country Name` 
             & dfm$variable %in% input$variable,]
  
  df1 <- df1 %>%
    group_by(iso3c) %>%
    arrange(date) %>%
    mutate(date = as.numeric(date)) %>% 
    ungroup() %>% 
    mutate(
      colors = rep(palette, ceil(nrow(.)/length(palette)))[1:nrow(.)]
    )

  
  texty <- paste("</br>", df1$`Short Name`,
                 "</br>Y: ", round(df1$value, 1), 
                 "</br>Year: ", df1$date)
  
  plot_ly(df1) %>%
    add_trace(
      x = ~ date,
      y = ~ value,
      name = ~ `Short Name`,
      type = 'scatter',
      mode = 'lines+markers',
      color = ~`Short Name`,
      colors = brewer.pal(12, "Paired"),
      hoverinfo = 'text',
      text = ~ texty
    ) %>%
    layout(legend = list(orientation = 'h'),
           xaxis = list(title = '', 
                        zeroline = FALSE, 
                        showline = FALSE, 
                        showgrid = TRUE
           ), 
           yaxis = list(title = str_wrap(input$variable, 40), 
                        zeroline = FALSE, 
                        showline = FALSE, 
                        showgrid = TRUE
           )
    )
  
})

```

COVID-19 Case Data
=====================================
  
Sidebar { .sidebar data-width=250}
-----------------------------------------------------------------------
Last Updated on: `r today`
  
```{r}

radioButtons("var",
             "",
             choices = covid_treemap,  
             selected = "COVID-19 Cases: Confirmed"
)  

# sliderInput("date",
#             "Day:",
#             min = min(treemap$date, na.rm = TRUE),
#             max = max(treemap$date, na.rm = TRUE),
#             value = max(treemap$date, na.rm = TRUE),
#             timeFormat="%Y-%m-%d")

getind <- reactive({
  input$var
})

```

Row
-----------------------------------------------------------------------
### `r renderText(getind())`
  
```{r}
renderPlotly({
  
  tmp <- treemap %>%
    filter(variable %in% input$var,
           date == max(date, na.rm = TRUE)) %>%
  #   filter(variable %in% input$var)
  # tmp$value[tmp$ids == "World"] <- tmp$value[tmp$ids == "World" & tmp$date == max(tmp$date, na.rm = TRUE)]
  # tmp <- tmp %>%
  #   filter(date == input$date) %>%
    left_join(color_map2, by = c("labels" = "Region")) 
  
  tmp %>% plot_ly(
    type ='treemap',
    ids =  ~ids,
    labels = ~labels,
    parents = ~parents,
    values = ~ value,
    branchvalues = "total",
    marker  = list(colors = ~color),
    domain = list(column=0))# %>% 
    # add_trace(
    #   branchvalues = "total"
    # )
  
})


```


### `r renderText(getind())`

```{r}

renderPlotly({
  
  dfb <- fullcd2 %>% 
    filter(date == max(date),
           variable %in% input$var) %>%
    arrange(desc(value))  %>%
    mutate(`Short Name` = fct_reorder(`Short Name`, desc(value))) %>% 
    slice(1:20)
  
  texty <- paste("</br>", dfb$`Short Name`,
                 "</br>Y: ", dfb$value, 
                 "</br>Date: ", dfb$date)
  
  plot_ly(dfb, 
          type = 'bar', 
          x = ~`Short Name`, 
          y = ~value, 
          color = ~Region,
          colors = reg_color,
          hoverinfo = 'text',
          text = ~ texty
  ) %>% 
    layout(
      legend = list(orientation = 'h', 
                    y = 100, 
                    x = 0),
      xaxis = list(title = '', 
                   zeroline = FALSE, 
                   showline = FALSE, 
                   showgrid = TRUE
      ), 
      yaxis = list(title = '', 
                   zeroline = FALSE, 
                   showline = FALSE, 
                   showgrid = TRUE
      )
    )
})


```

COVID-19 Severity
=====================================
Sidebar { .sidebar data-width=250}
-----------------------------------------------------------------------
Last Updated on: `r today`

```{r}

selectizeInput("countrysev",
               "Country:",
               c(unique(as.character(df$country))),
               multiple = TRUE, 
               options = list(maxItems = 4),
               selected = wst
)




radioButtons("var2",
             "Y-axis",
             choices = c("Country-specific", "Constant"),  
             selected = "Country-specific"
)  


```

Based on: <a href= "http://w0lxpjekins05.worldbank.org:3939/covid-dashboard-dev/#section-covid-19-severity">Schellekens, Philip and Sourrouille, Diego M., Covid-19 Mortality in Rich and Poor Countries: A Tale of Two Pandemics? (May 28, 2020). World Bank Policy Research Working Paper No. 9260.</a>


Row
-----------------------------------------------------------------------

### The severity ratio is defined as the ratio of deaths attributed to COVID-19 relative to expected total deaths without the pandemic in the same time period.
```{r}

#renderText("The severity ratio is defined as the ratio of deaths attributed to COVID-19 relative to expected total deaths without the pandemic in the same time period.")

####### Just trying chart out to observe data issues

set <- 1          #for displaying (first, second, etc set)
increment <- 4    #number of charts in each set

########## SUBSET ######################################################################################################################

uiOutput("dynamic2")

output$dynamic2 <- renderUI({ 
  if (input$var2 == "Country-specific"){
    shinycssloaders::withSpinner(
    ggiraphOutput("sev1", width = "100%", height = "800px")
    )
  } else {
    shinycssloaders::withSpinner(
    ggiraphOutput("sev2", width = "100%", height = "800px")
    )
    }
})

output$sev1 <- renderggiraph({
  
########## CHART  #######################################################################################################################
df$tooltip1 <- c(paste0("Daily Severity Ratio (%): ", round(100*(df$discrete_severity), 1)))
df$tooltip2 <- c(paste0("14-Day Average Severity Ratio (%): ", round(100*(df$cumulative_severity_14), 1)))
  
df2 <- df %>%
  subset(date>="2020-03-01") %>%
  subset(group==1) %>%
#  subset(country %in% c("Argentina", "Brazil", "Colombia", "India"))
  subset(country %in% input$countrysev)

#cols    <- c( "c1" = "#6baed6", "c2" = "black" )


p <- df2 %>%
  ggplot() +

#geoms--------------------------------------------------------------------------------------------------------

geom_segment_interactive( aes(x=date, xend=date, y=0, yend=discrete_severity*100), color="#6baed6", size=0.7, alpha=0.5, tooltip = df2$tooltip1) +
  geom_line_interactive(aes(x=date, y=cumulative_severity_14*100),color="black",size=0.3, alpha=0.7, tooltip = df2$tooltip2)+
  geom_line(aes(x=date, y=pmr_below*100), size=0.3,color="black", alpha=0.7)+
  geom_line(aes(x=date, y=pmr_above*100), size=0.3,color="black", alpha=0.7)+

#scales-----------------------------------------------------------------------------------------------------

#scale_x_date(date_labels = "%b%e" , date_breaks = "1 month", date_minor_breaks = "1 day", expand=c(0,0))+
#scale_y_continuous(limits = c(0, 50), expand=c(0,0))+

#labels-------------------------------------------------------------------------------------------------------
xlab("") +
  ylab("Daily severity ratio (percent)")+

  geom_label_repel(
    data=subset(df2, date=="2020-03-10"),
    force = 1,
    segment.size = 0.3,
    segment.color="grey50",
    segment.alpha=0.3,
    box.padding = 0.2,
    point.padding = 0,
    aes(x=date, y=pmr_below*100,label=cause_below),
    alpha=1,
    size=2.5) +

  geom_label_repel(
    data=subset(df2, date=="2020-03-10"),
    force = 1,
    segment.size = 0.3,
    segment.color="grey50",
    segment.alpha=0.3,
    box.padding = 0.2,
    point.padding = 0,
    aes(x=date, y=pmr_above*100,label=cause_above),
    alpha=1,
    size=2.5) +

#theme---------------------------------------------------------------------------------------------------

theme_light() +

  theme(
  #  legend.position="top",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title.y = element_text(color="black", size=7, vjust=2),
    axis.title.x = element_text(color="black", size=7, vjust=-2),
    axis.text =element_text(size=7,color="black"),
    legend.background=element_blank(),
  #  legend.title=element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    panel.spacing.x=unit(1, "lines"),panel.spacing.y=unit(1.6, "lines"),
    strip.text.x = element_text(size = 8, color = "black"),
    strip.text.y = element_text(size = 8, color = "black"),
    strip.background = element_rect(color="white", fill="white", size=1.5, 
                                    linetype="solid"))

#facet -----------------------------------------------------------------------------------------------------
p1 <- p + facet_wrap(~country, ncol=2, scales="free") +
      theme(strip.text.x = element_text(size = 11))

x <- girafe(ggobj = p1,
            width_svg = 18,
            height_svg = 10)

x

})


output$sev2 <- renderggiraph({
  
########## CHART  #######################################################################################################################
df$tooltip1 <- c(paste0("Daily Severity Ratio (%): ", round(100*(df$discrete_severity), 1)))
df$tooltip2 <- c(paste0("14-Day Average Severity Ratio (%): ", round(100*(df$cumulative_severity_14), 1)))
  
df2 <- df %>%
  subset(date>="2020-03-01") %>%
  subset(group==1) %>%
#  subset(country %in% c("Argentina", "Brazil", "Colombia", "India"))
  subset(country %in% input$countrysev)


p <- df2 %>%
  ggplot() +

#geoms--------------------------------------------------------------------------------------------------------

geom_segment_interactive( aes(x=date, xend=date, y=0, yend=discrete_severity*100), color="#6baed6",size=0.7, alpha=0.5, tooltip = df2$tooltip1) +
  geom_line_interactive(aes(x=date, y=cumulative_severity_14*100),color="black",size=0.3, alpha=0.7, tooltip = df2$tooltip2)+
  geom_line(aes(x=date, y=pmr_below*100), size=0.3,color="black", alpha=0.7)+
  geom_line(aes(x=date, y=pmr_above*100), size=0.3,color="black", alpha=0.7)+

#scales-----------------------------------------------------------------------------------------------------

#scale_x_date(date_labels = "%b%e" , date_breaks = "1 month", date_minor_breaks = "1 day", expand=c(0,0))+
#scale_y_continuous(limits = c(0, 50), expand=c(0,0))+

#labels-------------------------------------------------------------------------------------------------------
xlab("") +
  ylab("Daily severity ratio (percent)")+

  geom_label_repel(
    data=subset(df2, date=="2020-03-10"),
    force = 1,
    segment.size = 0.3,
    segment.color="grey50",
    segment.alpha=0.3,
    box.padding = 0.2,
    point.padding = 0,
    aes(x=date, y=pmr_below*100,label=cause_below),
    alpha=1,
    size=2.5) +

  geom_label_repel(
    data=subset(df2, date=="2020-03-10"),
    force = 1,
    segment.size = 0.3,
    segment.color="grey50",
    segment.alpha=0.3,
    box.padding = 0.2,
    point.padding = 0,
    aes(x=date, y=pmr_above*100,label=cause_above),
    alpha=1,
    size=2.5) +

#theme---------------------------------------------------------------------------------------------------

theme_light() +

  theme(
    legend.position="top",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title.y = element_text(color="black", size=7, vjust=2),
    axis.title.x = element_text(color="black", size=7, vjust=-2),
    axis.text =element_text(size=7,color="black"),
    legend.background=element_blank(),
    legend.title=element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    panel.spacing.x=unit(1, "lines"),panel.spacing.y=unit(1.6, "lines"),
    strip.text.x = element_text(size = 8, color = "black"),
    strip.text.y = element_text(size = 8, color = "black"),
    strip.background = element_rect(color="white", fill="white", size=1.5, 
                                    linetype="solid"))

#facet -----------------------------------------------------------------------------------------------------
p2 <- p + facet_wrap(~country, ncol=2) +
      theme(strip.text.x = element_text(size = 11))
  
#y <- girafe(ggobj = p1,
#          width_svg = 40, 
#          height_svg = 25)
  
y <- girafe(ggobj = p2,
            width_svg = 18,
            height_svg = 10)

y

})

```
