---
title: "Understanding the COVID-19 pandemic through data"
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    css: R-shiny-style-changes.css
---
  
```{r setup, include=FALSE}
source("styles.R")
source("00_setup.R")

library(pins)
library(RColorBrewer)
library(data.table)


my_key <- Sys.getenv("connect_key")
my_server <- "http://localhost:3939/"

# UNCOMMENT FOR LOCAL USE with EXTERNAL server:
# my_key <- Sys.getenv("connect_key_ext")
# my_server <- "http://w0lxopshyprd1b.worldbank.org:3939/" # External server

# UNCOMMENT FOR LOCAL USE with INTERNAL server:
# my_key <- Sys.getenv("connect_key_int")
# my_server <- "http://w0lxpjekins05.worldbank.org:3939/" # Internal server

board_register_rsconnect(key    = my_key,
                         server = my_server)

#--------- Data from server
dfm            <- pin_get("dfm", board = "rsconnect")
ccc            <- pin_get("ccc", board = "rsconnect")
fullcd2        <- pin_get("fullcd2", board = "rsconnect")
treemap        <- pin_get("treemap_data", board = "rsconnect")
wld_data       <- pin_get("wld_data", board = "rsconnect")
indicator_list <- pin_get("indicator_list", board = "rsconnect")
new_c          <- pin_get("new_c", board = "rsconnect")
new_wl         <- pin_get("new_wl", board = "rsconnect")
#cg_data       <- pin_get("cg_data", board = "rsconnect")
age_pop        <- pin_get("age_pop", board = "rsconnect")
age_pop        <- age_pop$age_pop
covid          <- pin_get("covid", board = "rsconnect")
health         <- pin_get("health", board = "rsconnect")
water          <- pin_get("water", board = "rsconnect")
df             <- pin_get("dfsev", board = "rsconnect")


#--------- covid data
covid          <- covid$covid
covid          <- c( "COVID-19 Cases: Case Fatality Rate",
                     "COVID-19 Cases: Confirmed",
                     "COVID-19 Cases: Confirmed (log scale)",
                     "COVID-19 Cases: Confirmed (per 1,000 people)",
                     "COVID-19 Cases: Confirmed (per 1,000 people, log scale)",
                     "COVID-19 Cases: Deaths",
                     "COVID-19 Cases: Deaths (log scale)",
                     "COVID-19 Cases: Deaths (per 1,000 people)",
                     "COVID-19 Cases: Deaths (per 1,000 people, log scale)",
                     "COVID-19 Cases: New Confirmed cases (Daily)",
                     "COVID-19 Cases: New Deaths (Daily)")
covidlog       <- covid[stringr::str_detect(covid, "log scale")]
covidbubble    <- covid[!stringr::str_detect(covid, "log scale")]
covid_treemap  <- covid[!stringr::str_detect(covid, paste(c("log scale", "Fatality", "per 1,000 people"),collapse = '|'))]


#--------- Aux data
health         <- health$health
water          <- water$water

my_date  <- max(wld_data$date)
my_date2 <- format(as.Date(my_date), format = "%b %d")
my_date2 <- as.character(my_date2)
today    <- as.Date(my_date) + 1

#--------- filtered cases
covid <- grep("(Confirmed|Deaths)", covid, value = TRUE)


#------country palette
#This has to be done here and not in utils because Ineed the `fullcd2` dataframe

uniq_cty <- unique(fullcd2$`Short Name`)

cty_color <-  setNames(
  rep(palette, ceiling(length(uniq_cty)/length(palette)))[1:length(uniq_cty)],
  uniq_cty
)


income <- c("Income level")

#--------- Find countries that are doing really bad

grcd <- fullcd2 %>% 
  select(`Short Name`, date, variable, value) %>% 
  filter(!is.na(`Short Name`)) %>% 
  as.data.table()


wst <- grcd[
  order(`Short Name`, variable, date)
     ][,
       # growth of each variable
       `:=`(gr = value/shift(value, 1, type = "lag") - 1,
            df = value - shift(value, 1, type = "lag") ),
       by = .(`Short Name`, variable)
     ][,
       # Rolling mean
       `:=`(rmg = frollmean(gr, 7, align =  "right"),  # rolling mean of growth
            rmd = frollmean(df, 7, align =  "right")), # rolling mean of diff
       by = .(`Short Name`, variable)
     ][
       # Select only one variable
       variable == "COVID-19 Cases: Deaths" & !is.na(rmg),
       # variable == "COVID-19 Cases: Confirmed",
       
       # select mrv of each group
       .SD[.N],
       by = .(`Short Name`)
     ][
       order(-value)
     ][
       1:20
     ][
       order(-rmd)  # Let's use the rolling mean of diff to include the US
     ][
       1:6,
       `Short Name`
     ]
rm(grcd)

df$date <- as.Date(df$date, format = "%Y-%m-%d")

```


Indicators
=====================================
  
Sidebar { .sidebar data-width=250}
-----------------------------------------------------------------------
Last Updated on: `r today`

```{r}

selectInput("variable",
            "Indicator Y-axis:",
            choices = list(
              `Age & Population` = age_pop,
              `COVID-19` = covid,
              Health = health,
              `Water & Sanitation` = water),  
            selected = "COVID-19 Cases: Deaths"
)    

selectInput("variable2",
            "Indicator X-axis:",
            choices = list(
              `Age & Population` = age_pop,
              `COVID-19` = covid,
              Health = health,
              `Water & Sanitation` = water),                
            selected = "Population ages 65 and above (% of total)"
) 

#selectInput("size",
#            "Bubble size:",
#            choices = covidbubble,                
#            selected = "COVID-19 Cases: Confirmed"
#) 

radioButtons("color",
             "Bubble color (click on legend items to exclude selected category):",
             choices = c("Income Group", "Region"),  
             selected = "Income Group"
) 

selectizeInput("Country Name",
               "Country (compare countries over time):",
               c(unique(as.character(dfm$`Short Name`))),
               multiple = TRUE, 
               options = list(maxItems = 12),
               selected = wst
)
```

Row  {data-height=130}
-----------------------------------------------------------------------
### World
```{r}
renderValueBox({
  con <- wld_data %>%
    filter(wld_data$date == max(wld_data$date)) %>%
    pull(`COVID-19 cases: Confirmed`) 
  con <- format(con, big.mark=",")
  de <- HTML(paste0("Total cases:<br/>", con))
  valueBox(tags$p(de, style = "font-size: 80%;")
           #, 
           #icon = "fa-users"
  )
})
```

### World
```{r}
renderValueBox({
  con <- new_wl %>% 
    filter(status == "Confirmed") %>% 
    pull(con) %>% 
    format(big.mark = ",")
  
  de <- HTML(paste0("New cases (", my_date2, "):<br/>", con))
  #de <- str_wrap(de, 15)
  valueBox(tags$p(de, style = "font-size: 80%;")
           #, 
           #icon = "fa-users"
  )
})
```

### World
```{r}
renderValueBox({
  con <- wld_data %>%
    filter(wld_data$date == max(wld_data$date)) %>%
    pull(`COVID-19 cases: Deaths`) %>%
    format(big.mark = ",")
  
  de <- HTML(paste0("Total deaths:<br/>", con))
  valueBox(tags$p(de, style = "font-size: 80%;")
           #, 
           #icon = "fa-users"
  )
})
```

Row  {data-height=500}
-----------------------------------------------------------------------
  
### Explore relationships
```{r}
renderPlotly ({
  df2 <- dfm[dfm$date == 2019 & dfm$variable %in% input$variable,]
  df3 <- dfm[dfm$date == 2019 & dfm$variable %in% input$variable2, c(1, 6)]
  dfs <- dfm[dfm$date == 2019 &
          dfm$variable %in% "COVID-19 Cases: Deaths", c(1, 6)]
  
  dfs <- dfs %>%
    rename("siz" = "mrv")
  
  df3 <- df3 %>%
    rename("mrv2" = "mrv")
  
  df4 <- merge(df2, df3, by = c("iso3c"))
  df4 <- merge(df4, dfs, by = c("iso3c"))
  
  m <- df4[df4$`Short Name` %in% input$`Country Name`, ]
  a <- list(
    x = m$mrv2,
    y = m$mrv,
    text = m$`Short Name`,
    xref = "x",
    yref = "y",
    showarrow = TRUE,
    arrowhead = 0,
    ax = 20,
    ay = -40,
    arrowcolor = toRGB("grey80"),
    font = list(color = "dark grey", size = 14)
  )
  
  
  texty <- paste("</br>",
                 df4$`Short Name`,
                 "</br>X: ",
                 round(df4$mrv2, 1),
                 "</br>Y: ",
                 round(df4$mrv, 1))
  
  
  df4$`Income Group` <- ordered(
    df4$`Income Group`,
    levels = c(
      "Low income",
      "Lower middle income",
      "Upper middle income",
      "High income"
    )
  )
  
  if (input$color == "Income Group") {
    df4$colort <- df4$`Income Group`
  } else {
    df4$colort <- df4$Region
  }
  
  if (input$color == "Income Group") {
    colorpal <- inc_color
  } else {
    colorpal <- brewer.pal(7, "Paired")
  }
  
  plot_ly(data = df4) %>%
    add_trace(
      x = ~ mrv2,
      y = ~ mrv,
      type = 'scatter',
      mode = 'markers',
      color =  ~ colort,
      colors = ~ colorpal,
      opacity = 0.9,
      size   = 15,
      #  sizes  = c(20, 100),
      hoverinfo = 'text',
      text = ~ texty
    ) %>%
    #add_annotations(text = df4$lab, textfont = list(size = 14),
    #         textposition = "top right") %>%
    layout(
      legend = list(
        x = 0,
        y = -0.3,
        orientation = 'h'
      ),
      annotations = a,
      xaxis = list(
        type = if_else(input$variable2 %in% covidlog, "log", "linear"),
        zeroline = FALSE,
        showline = FALSE,
        showgrid = TRUE,
        title = str_wrap((input$variable2), 40),
        range = c(min(df4$variable), max(df4$variable))
      ),
      yaxis = list(
        type = if_else(input$variable %in% covidlog, "log", "linear"),
        zeroline = FALSE,
        showline = FALSE,
        showgrid = TRUE,
        title = str_wrap(input$variable, 40),
        range = c(min(df4$variable), max(df4$variable))
      )
    )
    }
) # end of renderPlotly
```


### Compare countries over time
```{r}

uiOutput("dynamic")

output$dynamic <- renderUI({ 
  if (input$variable %in% covid){
    plotlyOutput("plot1")
  } else {
    plotlyOutput("plot2")
  }
})

output$plot1 <- renderPlotly({
  
  dft <- fullcd2[fullcd2$`Short Name` %in% input$`Country Name`
                 & fullcd2$variable %in% input$variable,]
  
  dft <- dft         %>%
    arrange(date)    %>% 
    group_by(Region) %>% 
    mutate(
      op = match(name, unique(name)), 
      opacity = (1-(op/max(op))+1/max(op)),
    ) %>% 
    ungroup() %>% 
    mutate(
      colors = rep(palette, ceil(nrow(.)/length(palette)))[1:nrow(.)]
    )  
  
  
  texty <- paste("</br>", dft$`Short Name`,
                 "</br>Y: ", round(dft$value, 2), 
                 "</br>Date: ", dft$date)
  
  plot_ly(data = dft,
          x    = ~date,
          y    = ~ value ) %>%
    add_trace(

      name      = ~ `Short Name`, 
      type      = 'scatter',
      mode      = 'lines+markers', 
      color     = ~`Short Name`,
      colors    = brewer.pal(12, "Paired"),
      hoverinfo = 'text',
      text      = ~ texty
    ) %>%
    layout(
      legend     = list(orientation = 'h'),
      xaxis      = list(
        title    = '',
        zeroline = FALSE,
        showline = FALSE,
        showgrid = TRUE
      ),
      yaxis      = list(type = if_else(input$variable %in% covidlog, "log", "linear"),
        title    = input$variable,
        zeroline = FALSE,
        showline = FALSE,
        showgrid = TRUE
      )
    )
  
})

output$plot2 <- renderPlotly({
  
  df1 <- dfm[dfm$`Short Name` %in% input$`Country Name` 
             & dfm$variable %in% input$variable,]
  
  df1 <- df1 %>%
    group_by(iso3c) %>%
    arrange(date) %>%
    mutate(date = as.numeric(date)) %>% 
    ungroup() %>% 
    mutate(
      colors = rep(palette, ceil(nrow(.)/length(palette)))[1:nrow(.)]
    )

  
  texty <- paste("</br>", df1$`Short Name`,
                 "</br>Y: ", round(df1$value, 1), 
                 "</br>Year: ", df1$date)
  
  plot_ly(df1) %>%
    add_trace(
      x = ~ date,
      y = ~ value,
      name = ~ `Short Name`,
      type = 'scatter',
      mode = 'lines+markers',
      color = ~`Short Name`,
      colors = brewer.pal(12, "Paired"),
      hoverinfo = 'text',
      text = ~ texty
    ) %>%
    layout(legend = list(orientation = 'h'),
           xaxis = list(title = '', 
                        zeroline = FALSE, 
                        showline = FALSE, 
                        showgrid = TRUE
           ), 
           yaxis = list(title = str_wrap(input$variable, 40), 
                        zeroline = FALSE, 
                        showline = FALSE, 
                        showgrid = TRUE
           )
    )
  
})

```

COVID-19 Case Data
=====================================
  
Sidebar { .sidebar data-width=250}
-----------------------------------------------------------------------
Last Updated on: `r today`
  
```{r}

radioButtons("var",
             "",
             choices = covid_treemap,  
             selected = "COVID-19 Cases: Confirmed"
)  

# sliderInput("date",
#             "Day:",
#             min = min(treemap$date, na.rm = TRUE),
#             max = max(treemap$date, na.rm = TRUE),
#             value = max(treemap$date, na.rm = TRUE),
#             timeFormat="%Y-%m-%d")

getind <- reactive({
  input$var
})

```

Row
-----------------------------------------------------------------------
### `r renderText(getind())`
  
```{r}
renderPlotly({
  
  tmp <- treemap %>%
    filter(variable %in% input$var,
           date == max(date, na.rm = TRUE)) %>%
  #   filter(variable %in% input$var)
  # tmp$value[tmp$ids == "World"] <- tmp$value[tmp$ids == "World" & tmp$date == max(tmp$date, na.rm = TRUE)]
  # tmp <- tmp %>%
  #   filter(date == input$date) %>%
    left_join(color_map2, by = c("labels" = "Region")) 
  
  tmp %>% plot_ly(
    type ='treemap',
    ids =  ~ids,
    labels = ~labels,
    parents = ~parents,
    values = ~ value,
    branchvalues = "total",
    marker  = list(colors = ~color),
    domain = list(column=0))# %>% 
    # add_trace(
    #   branchvalues = "total"
    # )
  
})


```


### `r renderText(getind())`

```{r}

renderPlotly({
  
  dfb <- fullcd2 %>% 
    filter(date == max(date),
           variable %in% input$var) %>%
    arrange(desc(value))  %>%
    mutate(`Short Name` = fct_reorder(`Short Name`, desc(value))) %>% 
    slice(1:20)
  
  texty <- paste("</br>", dfb$`Short Name`,
                 "</br>Y: ", dfb$value, 
                 "</br>Date: ", dfb$date)
  
  plot_ly(dfb, 
          type = 'bar', 
          x = ~`Short Name`, 
          y = ~value, 
          color = ~Region,
          colors = reg_color,
          hoverinfo = 'text',
          text = ~ texty
  ) %>% 
    layout(
      legend = list(orientation = 'h', 
                    y = 100, 
                    x = 0),
      xaxis = list(title = '', 
                   zeroline = FALSE, 
                   showline = FALSE, 
                   showgrid = TRUE
      ), 
      yaxis = list(title = '', 
                   zeroline = FALSE, 
                   showline = FALSE, 
                   showgrid = TRUE
      )
    )
})


```

COVID-19 vs other causes of death
=====================================
Sidebar { .sidebar data-width=250}
-----------------------------------------------------------------------
Last Updated on: `r today`

```{r}

selectizeInput("unit_selected",
               "Country:",
               c(unique(as.character(df$country))),
               multiple = TRUE, 
               options = list(maxItems = 4),
               selected = wst
)

dateInput("start_date",
          "Start Date:",
          value = min(df$date),
          min = min(df$date),
          max = max(df$date)
          #,
#          end = max(df$date)
)

radioButtons("linevar",
             "Cause of death closest to",
          #   choiceNames  = c("Peak value", "Most recent value"),  
          #   choiceValues = c("peak_pmr", "last_pmr"),
             choices = c("Peak value", "Last value"),  
             selected = "Peak value"
) 

radioButtons("var2",
             "Y-axis",
             choices = c("Country-specific", "Constant"),  
             selected = "Country-specific"
)  


```

Source: <a href= "https://documents.worldbank.org/en/publication/documents-reports/documentdetail/559181590712052524/covid-19-mortality-in-rich-and-poor-countries-a-tale-of-two-pandemics" target="_blank"> Schellekens and Sourrouille (World Bank Working Paper, 2020)</a>, <a href= "http://ghdx.healthdata.org/gbd-2019" target="_blank"> Global Burden of Disease Study 2019 (IHME, 2020)</a> and <a href= "https://coronavirus.jhu.edu/" target="_blank"> JHU CSSE COVID-19 data</a>.

Row
-----------------------------------------------------------------------

### COVID-19 severity ratio expresses COVID-19 deaths in 2020 to deaths from all causes in 2019 as 7-day trailing average over the same period. Top-cause severity ratio(s) show cause-specific deaths in 2019 to deaths from all causes in 2019 as yearly average. Top cause(s) in 2019 are chosen from 133 level-3 disease families of the 2019 Global Burden of Disease Study so that its ratio approximates the peak and most recent 7-day COVID-19 ratio most closely from below.
```{r}

####### Just trying chart out to observe data issues

set <- 1          #for displaying (first, second, etc set)
increment <- 4    #number of charts in each set

########## SUBSET ######################################################################################################################

uiOutput("dynamic2")

output$dynamic2 <- renderUI({ 
  if (input$var2 == "Country-specific"){
    shinycssloaders::withSpinner(
    ggiraphOutput("sev1", width = "100%", height = "600px")
    )
  } else {
    shinycssloaders::withSpinner(
    ggiraphOutput("sev2", width = "100%", height = "600px")
    )
    }
})

output$sev1 <- renderggiraph({
  
########## CHART  #######################################################################################################################
#group_selected <- 1
#unit_selected <- "Belgium"
#start_date <- "2020-03-01"

# Prepare dataframe
df$date <- as.Date(df$date, format = "%Y-%m-%d")
#df$tooltip1 <- c(paste0("Daily Severity Ratio (%): ", round(100*(df$discrete_severity), 1)))
#start_date <- min(df$date)

df_chart <- df %>%       # Need to pick date range before calculating what follows
  subset(date>=input$start_date) %>%
#  subset(date>=start_date) %>%
#  subset(country %in% unit_selected)
#  subset(group==group_selected) %>%
  subset(country %in% input$unit_selected)

df_chart <- df_chart %>%
  group_by(country) %>%
  mutate(peak_pmr = case_when(is.na(peak_above_pmr) ~ peak_below_pmr,    # We would always like to select the cause of death that is just below the reference value. In some exceptional cases the reference value is so low that there is no threshold below. We will then pick the nearest threshold above.
                              is.na(peak_below_pmr) ~ peak_above_pmr,
                              TRUE ~ peak_below_pmr),
         peak_rank = case_when(is.na(peak_above_rank) ~ peak_below_rank,
                               is.na(peak_below_rank) ~ peak_above_rank,
                               TRUE ~ peak_below_rank),
         peak_name = case_when(is.na(peak_above_name) ~ peak_below_name,
                               is.na(peak_below_name) ~ peak_above_name,
                               TRUE ~ peak_below_name), 
         last_pmr = case_when(is.na(last_above_pmr) ~ last_below_pmr,
                              is.na(last_below_pmr) ~ last_above_pmr,
                              TRUE ~ last_below_pmr),
         last_rank = case_when(is.na(last_above_rank) ~ last_below_rank,
                               is.na(last_below_rank) ~ last_above_rank,
                               TRUE ~ last_below_rank),
         last_name = case_when(is.na(last_above_name) ~ last_below_name,
                               is.na(last_below_name) ~ last_above_name,
                               TRUE ~ last_below_name)) 

df_chart$labelcause <- ifelse(df_chart$peak_rank==df_chart$last_rank, 
          # allowing for the possibility that values approximating peak and last are the same
                        ifelse(df_chart$peak_rank==0,               
          # the 0 rank corresponds to all-cause mortality in 2019
                               "All Causes", 
                               paste0(df_chart$country2, ": Cause #", df_chart$peak_rank, " = ", df_chart$peak_name)),
                        ifelse(df_chart$peak_rank==0 ,
                               paste0(df_chart$country2, ": All Causes and Cause #", df_chart$last_rank, " = ", df_chart$last_name), 
                               paste0(df_chart$country2, ": Cause #", df_chart$peak_rank, " = ", df_chart$peak_name, "; Cause #", df_chart$last_rank, " = ", df_chart$last_name)
                               ))


# chart
df_chart$labelcause <- str_wrap(df_chart$labelcause, 60)
df_chart$cumsev7 <- round(100*(df_chart$cumulative_severity_7),1)
df_chart$tooltip2 <- paste0("7-Day Average Severity Ratio: ", df_chart$cumsev7, "%")

top <- plyr::round_any(df_chart$peak_7[1], ifelse(df_chart$peak_7[1]<0.1, 0.01, 0.1), ceiling)  # max range of display on Y axis

df_chart <- df_chart %>%
  group_by(country) %>%
  mutate(chosen_pmr = ifelse(input$linevar == "Peak value", peak_pmr, last_pmr)) %>%
  mutate(chosen_rank = ifelse(input$linevar == "Peak value", peak_rank, last_rank))

#df_chart$chosen_rank <- ifelse(input$linevar == "Peak value", df_chart$peak_rank,
#                              df_chart$last_rank)

p <- df_chart %>% 
  ggplot() +
  geom_line(aes(x=date, y=cumulative_severity_7),       # show the 7-day average
            size=1.5, color="#619CFF", alpha=1)+
    geom_point_interactive(aes(x=date, y=cumulative_severity_7, tooltip = tooltip2),       # show the 7-day average
            size=0.5, color="#619CFF", alpha=1)+
  geom_line(aes(x=date, y=chosen_pmr),                    # show the horizontal line for the cause of death that is closest from below to the peak value of the 7-day average
            size=1.5, color="#F8766D", alpha=1)+
#  geom_line(aes(x=date, y=last_pmr),                    # show the horizontal line for the cause of death that is closest from below to the last value of the 7-day average
#            size=1.5,color="#F8766D", alpha=1)+
  
  scale_x_date(date_breaks = "1 month", date_labels = "%b")+
  scale_y_continuous(#limits = c(NA, top), 
                     expand = expansion(mult = c(0.01, 0)),
                     labels = scales::percent_format(accuracy = ifelse(top<0.1, 0.1, 1))) +  #show full percentages except when peak value is less than 10%.
  
  labs(x="", 
       y="Percent of 2019 deaths",
       title=paste0("Severity Ratios for COVID-19 and Top Cause(s) of Death in 2019"))+
  
  geom_label_repel(                                              # Label for peak cause
    data=subset(df_chart, date=="2020-08-01"),
    force = 1,
    point.padding = 0, label.padding = 0.5,
    aes(x=date, y=chosen_pmr,
        label=ifelse(chosen_rank==0, "All Causes", paste0("Cause #", chosen_rank))),
    color="#F8766D",
    label.size=1.2,
    hjust=0.5, alpha=1,size=5) +
  
  geom_label_repel(                                            # Labe for COVID-19
    data=subset(df_chart, date==max(date)),
    force = 1,
    point.padding = 0,label.padding = 0.5,
    nudge_x = -10,
    aes(x=date, y=cumulative_severity_7,label="COVID-19"),
    color="#619CFF",
    label.size=1.2,
    hjust=0, alpha=1, size=5) +
  
  theme_light() +
  theme(
    legend.position="none",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title.y = element_text(color="black", size=14, vjust=4),
    axis.title.x = element_text(color="black", size=14, vjust=0),
    axis.text =element_text(size=14,color="black"),
    legend.background=element_blank(),
    legend.title=element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    panel.spacing.x=unit(2, "lines"),panel.spacing.y=unit(2, "lines"),
    strip.text.x = element_text(size = 9, color = "black", vjust=1),
    strip.text.y = element_text(size = 9, color = "black", vjust=1),
    strip.background = element_rect(color="white", fill="white", size=1.5, linetype="solid"),
    plot.margin = margin(2, 3, 1, 1.5, "cm"),
    plot.title=element_text(size=21, vjust=7,hjust=0.5),
    plot.subtitle = element_text(size=16, vjust=5, hjust=0.5),
    plot.caption = element_text(size=12, hjust=0, vjust=0, color="darkgrey")
  )+
  coord_cartesian(clip = 'off')

# #facet -----------------------------------------------------------------------------------------------------
 p1 <- p + facet_wrap(~labelcause, ncol=2, scales="free") +
       theme(strip.text.x = element_text(size = 20))

 x <- girafe(ggobj = p1,
             width_svg = 18,
             height_svg = 10)
 
 x
# 

})
# 
# 
output$sev2 <- renderggiraph({
  
########## CHART  #######################################################################################################################
#group_selected <- 1
#unit_selected <- "Belgium"
#start_date <- "2020-03-01"

# Prepare dataframe
df$date <- as.Date(df$date, format = "%Y-%m-%d")
#df$tooltip1 <- c(paste0("Daily Severity Ratio (%): ", round(100*(df$discrete_severity), 1)))
#start_date <- min(df$date)

df_chart <- df %>%       # Need to pick date range before calculating what follows
  subset(date>=input$start_date) %>%
#  subset(date>=start_date) %>%
#  subset(country %in% unit_selected)
#  subset(group==group_selected) %>%
  subset(country %in% input$unit_selected)

df_chart <- df_chart %>%
  group_by(country) %>%
  mutate(peak_pmr = case_when(is.na(peak_above_pmr) ~ peak_below_pmr,    # We would always like to select the cause of death that is just below the reference value. In some exceptional cases the reference value is so low that there is no threshold below. We will then pick the nearest threshold above.
                              is.na(peak_below_pmr) ~ peak_above_pmr,
                              TRUE ~ peak_below_pmr),
         peak_rank = case_when(is.na(peak_above_rank) ~ peak_below_rank,
                               is.na(peak_below_rank) ~ peak_above_rank,
                               TRUE ~ peak_below_rank),
         peak_name = case_when(is.na(peak_above_name) ~ peak_below_name,
                               is.na(peak_below_name) ~ peak_above_name,
                               TRUE ~ peak_below_name), 
         last_pmr = case_when(is.na(last_above_pmr) ~ last_below_pmr,
                              is.na(last_below_pmr) ~ last_above_pmr,
                              TRUE ~ last_below_pmr),
         last_rank = case_when(is.na(last_above_rank) ~ last_below_rank,
                               is.na(last_below_rank) ~ last_above_rank,
                               TRUE ~ last_below_rank),
         last_name = case_when(is.na(last_above_name) ~ last_below_name,
                               is.na(last_below_name) ~ last_above_name,
                               TRUE ~ last_below_name)) 

df_chart$labelcause <- ifelse(df_chart$peak_rank==df_chart$last_rank, 
          # allowing for the possibility that values approximating peak and last are the same
                        ifelse(df_chart$peak_rank==0,               
          # the 0 rank corresponds to all-cause mortality in 2019
                               "All Causes", 
                               paste0(df_chart$country2, ": Cause #", df_chart$peak_rank, " = ", df_chart$peak_name)),
                        ifelse(df_chart$peak_rank==0 ,
                               paste0(df_chart$country2, ": All Causes and Cause #", df_chart$last_rank, " = ", df_chart$last_name), 
                               paste0(df_chart$country2, ": Cause #", df_chart$peak_rank, " = ", df_chart$peak_name, "; Cause #", df_chart$last_rank, " = ", df_chart$last_name)
                               ))


# chart
df_chart$labelcause <- str_wrap(df_chart$labelcause, 60)
df_chart$cumsev7 <- round(100*(df_chart$cumulative_severity_7),1)
df_chart$tooltip2 <- paste0("7-Day Average Severity Ratio: ", df_chart$cumsev7, "%")

top <- plyr::round_any(df_chart$peak_7[1], ifelse(df_chart$peak_7[1]<0.1, 0.01, 0.1), ceiling)  # max range of display on Y axis

df_chart$chosen_pmr <- ifelse(input$linevar == "Peak value", df_chart$peak_pmr,
                              df_chart$last_pmr)

df_chart$chosen_rank <- ifelse(input$linevar == "Peak value", df_chart$peak_rank,
                              df_chart$last_rank)

p <- df_chart %>% 
  ggplot() +
  geom_line(aes(x=date, y=cumulative_severity_7),       # show the 7-day average
            size=1.5, color="#619CFF", alpha=1)+
    geom_point_interactive(aes(x=date, y=cumulative_severity_7, tooltip = tooltip2),       # show the 7-day average
            size=0.5, color="#619CFF", alpha=1)+
  geom_line(aes(x=date, y=chosen_pmr),                    # show the horizontal line for the cause of death that is closest from below to the peak value of the 7-day average
            size=1.5, color="#F8766D", alpha=1)+
#  geom_line(aes(x=date, y=last_pmr),                    # show the horizontal line for the cause of death that is closest from below to the last value of the 7-day average
#            size=1.5,color="#F8766D", alpha=1)+
  
  scale_x_date(date_breaks = "1 month", date_labels = "%b")+
  scale_y_continuous(#limits = c(NA, top), 
                     expand = expansion(mult = c(0.01, 0)),
                     labels = scales::percent_format(accuracy = ifelse(top<0.1, 0.1, 1))) +  #show full percentages except when peak value is less than 10%.
  
  labs(x="", 
       y="Percent of 2019 deaths",
       title=paste0("Severity Ratios for COVID-19 and Top Cause(s) of Death in 2019"))+
  
  geom_label_repel(                                              # Label for peak cause
    data=subset(df_chart, date==min(date)),
    force = 1,
    point.padding = 0, label.padding = 0.5,
    aes(x=date, y=chosen_pmr,
        label=ifelse(chosen_rank==0, "All Causes", paste0("Cause #", chosen_rank))),
    color="#F8766D",
    label.size=1.2,
    hjust=0.5, alpha=1,size=5) +
  
  geom_label_repel(                                            # Labe for COVID-19
    data=subset(df_chart, date==max(date)),
    force = 1,
    point.padding = 0,label.padding = 0.5,
    nudge_x = -10,
    aes(x=date, y=cumulative_severity_7,label="COVID-19"),
    color="#619CFF",
    label.size=1.2,
    hjust=0, alpha=1, size=5) +
  
  theme_light() +
  theme(
    legend.position="none",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title.y = element_text(color="black", size=14, vjust=4),
    axis.title.x = element_text(color="black", size=14, vjust=0),
    axis.text =element_text(size=14,color="black"),
    legend.background=element_blank(),
    legend.title=element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    panel.spacing.x=unit(2, "lines"),panel.spacing.y=unit(2, "lines"),
    strip.text.x = element_text(size = 9, color = "black", vjust=1),
    strip.text.y = element_text(size = 9, color = "black", vjust=1),
    strip.background = element_rect(color="white", fill="white", size=1.5, linetype="solid"),
    plot.margin = margin(2, 3, 1, 1.5, "cm"),
    plot.title=element_text(size=21, vjust=7,hjust=0.5),
    plot.subtitle = element_text(size=16, vjust=5, hjust=0.5),
    plot.caption = element_text(size=12, hjust=0, vjust=0, color="darkgrey")
  )+
  coord_cartesian(clip = 'off')

# #facet -----------------------------------------------------------------------------------------------------
 p2 <- p + facet_wrap(~labelcause, ncol=2, scales="fixed") +
       theme(strip.text.x = element_text(size = 20))

 y <- girafe(ggobj = p2,
             width_svg = 18,
             height_svg = 10)
 
 y
# 

})
```
